%%%%%%%%%%%%%%%%%
% Main document.
%%%%%%%%%%%%%%%%%

\documentclass[11pt]{book}
\usepackage{cacdmanual}
% \usepackage{alltt}
% \usepackage{fancyhdr}
% \usepackage{graphics}
% \usepackage{calc}

\usepackage{a4wide,times,latexsym,calc,ifthen,float}
\usepackage{epsfig}
\usepackage{fancyvrb}
\usepackage[dvips,colorlinks]{hyperref}

\newcommand{\io}[1]{{\ttfamily #1}}
\newcommand{\CACDTOP}[1]{\io{\$ICDPATH/share/#1}}
\newcommand{\manualpage}[1]{\io{icdman\ #1}}
\input{list.sty}

\def\floatpagefraction{0.9}
\def\textfraction{0.1}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%
% Define title page.
%%%%%%%%%%%%%%%%%%%%%

\cacdreport{ET-ENS 2006.01}
\cacdauthors{Nick van der Meijs, Simon de Graaf}
\cacddate{June 29, 2006}
\cacdyears{2006}
\cacdtitle{SPACE EXAMPLES}
\cacdheadertitle{Space Examples}
\cacdmaketitle

\tableofcontents

%%%%%%%%%%%%%%%%%%%%
% Define Remainder.
%%%%%%%%%%%%%%%%%%%%

\chapter{Introduction}

This document explains something about the space demo examples.
Trying these demo's
can be a good starting point to learn more about the Space System.

Before you can use the demo's, you must know the space software installation path.
When the space software is installed in your home directory, then the installation
path (or \io{ICDPATH}) is \io{\$HOME/cacd}.
But you can also rename 'cacd' into something else, for example 'cacd\_new'.
The only thing you need to do, is to set your \io{PATH} environment variable to the \io{ICDPATH}.
See also the file \io{\$ICDPATH/Installation.txt} for more information.
Give for example the following commands to set your \io{PATH}:
\begin{Verbatim}
    % setenv ICDPATH $HOME/cacd
    % setenv PATH $ICDPATH/bin:$PATH
\end{Verbatim}
Go now to the demo directory, for example type:
\begin{Verbatim}
    % cd $ICDPATH/share/demo
    % ls
    attenua  invert      ny9t    README    sub3term   switchbox
    crand    multiplier  poly5   sram      suboscil
\end{Verbatim}
Each demo contains a \io{README} file, which explains what you must do.
When you don't have a private software installation or don't have write permission,
you must copy the demo files to a private directory to do the demo.
For example:
\begin{Verbatim}
    % mkdir mydemo ; cd mydemo
    % cp $ICDPATH/share/demo/attenua/* .
    % cat README
\end{Verbatim}
When you don't want to type the space commands yourself,
you can use a shell script to do it for you.
In that case, type:
\begin{Verbatim}
    % ./script.sh
 or
    % sh script.sh
\end{Verbatim}
Read the following chapters, to know more about the demo examples (if included).
We wish you success.
We hope that you appreciate our software and want to use it.
If you have any questions or comments, they are welcome.
\begin{Verbatim}
    The SPACE team.
\end{Verbatim}

\include{attenua/demo}
\include{crand/demo}
\include{invert/demo}
\include{multiplier/demo}
\include{ny9t/demo}
\include{poly5/demo}
\include{sram/demo}
\include{sub3term/demo}
\include{suboscil/demo}
\include{switchbox/demo}

% \appendix

\end{document}
