% -*- latex -*-

\chapter{Acknowledgments}
%\pagestyle{fancy}
All software and design effort for the entire 'fishbone' Sea-of-Gates design
trajectory (figure~\ref{oceantools}) was performed 'in house' at Delft
University of Technology, department of Electrical Engineering. People from
various groups within Electrical Engineering participated, making it a
faculty-wide project. The people who contibuted noticably to the creation of
the ocean system are listed below:

\begin{description}
\item[~]
\begin{itemize}
% ? doesn't work \addtolength{\leftmargin}{2.5cm}
\item
{\sc Software}
\begin{itemize}
\item[Circuit tools 'sls', 'simeye':]
Arjan van Genderen 
\item[Extractor 'space':]
Nick van der Meijs and Simon de Graaf
\item[Ocean placer 'Madonna':]
Paul Stravers and Irek Karkowski
\item[Ocean router 'trout':]
Patrick Groeneveld
\item[Layout editor 'seadali':]
Patrick Groeneveld (based on 'dali' by Pieter van der Wolf)
\item[Purifiers 'fish'and 'ghoti':]
Patrick Groeneveld and Paul Stravers
\item[Port to {\sc LINUX} and sun4:]
Paul Stravers and Irek Karkowski
\item[Database support (seadif):]
Paul Stravers and Patrick Groeneveld
\item[Database support (Nelsis3):]
Nick van der Meijs, Simon de Graaf and Patrick Groeneveld
\item[Database support (Nelsis4):]
Alfred van der Hoeven and Pieter van der Wolf
\item[Nelsis backbone system:]
Peter Bingley, Hong Cai, Arjan van Genderen, Patrick Groeneveld, Alfred van der
Hoeven, Fons de Lange, Ren\'e van Leuken, Nick van der Meijs, Paul Stravers and
Pieter van der Wolf.
\item[Overall software coordination:]
Patrick Groeneveld 
\end{itemize}
\item
{\sc The Chip}
\begin{itemize}
\item[The 'fishbone' image:]
Paul Stravers and Patrick Groeneveld
\item[Multi-project chip:]
Paul Stravers, Patrick Groeneveld, Anton Bakker, Antoon Frehe
\item[Library design:]
Anton Bakker, Emile Hendriks, Peter van de Reest, Paul Stravers and Patrick
Groeneveld
\item[Design chip:]
Reinder Nouta, Paul Stravers, and Antoon Frehe 
\item[Processing empty wafers:]
Philips Research Laboratories (via Reinder Nouta)
\item[Metalization (DIMES):]
Hugo Schellevis, prof. Cees Beenakker 
\item[Chip testing:]
Rob Bruinink and Hubert Schuit
\end{itemize}
\item
{\sc Documentation}
\begin{itemize}
\item[Manuals/tutorials:]
Paul Stravers, Patrick Groeneveld, Nick van der Meijs, Arjan van Genderen, Simon de Graaf,
Anton Bakker, Emile Hendriks, and Reinder Nouta.
\end{itemize}
\item
{\sc And finally}
\begin{itemize}
\item[The OP practicum commission:]
Emile Hendriks, Anton Bakker, Reinder Nouta,
Chris Verhoeven,\\
Cees Beenakker and Patrick Groeneveld
\item[Financial support by:]
Faculty of Electrical Engineering, TU Delft\\
PICO (wafer processing, test equipment)
\item[Related projects:]
IOP Sea-of-Gates project \\
Royal Dutch Academy of Arts and Sciences
\item[Special thanks to:]
Prof. Ralph Otten, Prof. Patrick Dewilde and many others.
\end{itemize}
\end{itemize}
\end{description}
